---
title: "cleaning for template data"
author: "Sascha Siegmund"
date: "`r Sys.Date()`"
output: 
  github_document:
    fig_width: 10
    fig_height: 7
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "nb_figs/clean_",
  warning = FALSE,
  message = FALSE
)
```



## purpose of notebook

  - [ ] initial look at data to get a basic understanding and gather todos -> df_raw
  - [ ] tidy up and cleaning of data set -> df
  - [ ] all processing of the data set happens here and is saved to pre-/processed csv-> df
 

## observations

  - 
               

## load packages

```{r load packages, setup, message=FALSE}
library(tidyverse) # tidy data frame
library(lubridate) # functions to work with date-times and time-spans
library(janitor) # expedite the initial data exploration and cleaning that comes with any new data set
library(patchwork) # separate ggplots into the same graphic
library(readxl) # read_excel to read xslx dataset
```


## import data

```{r}
df_raw <- read_csv('../data/house_prices_regression/train.csv')
df <- df_raw
```


## first look at data

```{r}
head(df_raw, 25)
```
```{r}
tail(df_raw)
```
```{r}
summary(df_raw)
```


## missing values

-  

```{r fig.height=20, fig.width=14}
#create data frame with information on whether the value in each cell is zero
tmp_df <- df_raw == 0
missing_by_column <- tmp_df %>%
  as_tibble %>% # convert to data frame
  mutate(row_number = 1:nrow(.)) %>% # add a column with the row number
  gather(variable, is_missing, -row_number) # turn wide data into narrow data

# plot the missing values in our data frame
ggplot(missing_by_column, aes(x = variable, y = row_number, fill =  is_missing)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_grey(name = "",
                  labels = c("present", "is zero", "is NA")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10)) +
  scale_y_reverse() +
  labs(x = "variables in dataset",
       y = "rows / observations")
```


## duplicated rows

- 

```{r fig.height=14, fig.width=10}
df_raw %>% get_dupes() # examine duplicate rows

# get row number of duplicated rows
tmp_df <- distinct(df_raw)
duplicated_rows <- as_tibble(rownames(df_raw)) %>% 
  mutate(duplicated = value %in% setdiff(rownames(df_raw), rownames(tmp_df))) %>% 
  mutate(value = as.numeric(value))

# plot duplicated rows as black lines
ggplot(duplicated_rows, aes(y = value, x = 1, fill = duplicated)) +
  geom_tile() + # plot a black line for each duplicated row
  ggtitle("row index of duplicated rows") + # add title
  theme_minimal() +
  scale_fill_grey(start=0.8, end=0.2,
                  name = "",
                  labels = c("ok", "duplicated")) +
  scale_y_reverse() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(y = "row index",
       x = "")
```


## constant columns

- 

```{r fig.width=14}
# get column name of constant columns
tmp_df <- remove_constant(df_raw, na.rm = TRUE)
constant_columns <- as_tibble(names(df_raw)) %>% 
  mutate(constant = value %in% setdiff(names(df_raw), names(tmp_df)))

# plot constant columns as black lines
ggplot(constant_columns, aes(y = value, x = 1, fill = constant)) +
  geom_tile() + # plot a black line for each constant column
  ggtitle("names of constant columns") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() + # reverse x axis
  theme_minimal() +
  scale_fill_grey(start=0.8, end=0.2,
                  name = "",
                  labels = c("ok", "constant")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  labs(y = "variables in dataset",
       x = "")
```


## near zero variance columns

- 

```{r fig.width=14}
# get column name and info of near zero variance column
nearZero_info <- caret::nearZeroVar(df_raw, freqCut = 95/5, uniqueCut = 10, saveMetrics = TRUE) %>%
  rownames_to_column(var = "name")
nearZero_info

# plot near zero variance columns as black lines
ggplot(nearZero_info, aes(y = name, x = 1, fill = nzv)) +
  geom_tile() + # plot a black line for each near zero var column
  ggtitle("names of near zero variance columns") + # add title
  coord_flip() + # flip x and y axis
  scale_x_reverse() + # reverse x axis
  theme_minimal() +
  scale_fill_grey(start=0.8, end=0.2,
                  name = "",
                  labels = c("ok", "near zero")) +
  theme(axis.text.x = element_text(angle=45, vjust=0.7, size=10),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) +
  labs(y = "variables in dataset",
       x = "")

# plot infos from nearZeroVar
p1 <- nearZero_info %>% ggplot(aes(x = freqRatio, fill = nzv)) +   geom_histogram(bins = 100) + 
  theme_minimal() + theme(legend.position="none")
p2 <- nearZero_info %>% ggplot(aes(x = percentUnique, fill = nzv)) + geom_histogram(bins = 100) + 
  theme_minimal() + theme(legend.position="none")
p3 <- nearZero_info %>% ggplot(aes(x = zeroVar)) + geom_bar() + 
  theme_minimal() + theme(legend.position="none")
p4 <- nearZero_info %>% ggplot(aes(x = nzv, fill = nzv)) + geom_bar() + 
  theme_minimal() + theme(legend.position="none") + labs(x = "near zero variance")
p5 <- nearZero_info %>% ggplot(aes(x = freqRatio, y = percentUnique, color = nzv)) + geom_point() + 
  theme_minimal() + theme(legend.position="none")

p5 / (p1 / p2 / (p3 | p4))
```


## cleaning

- 

```{r}
# df <- df %>%
#   clean_names() %>%  # clean column names
#   distinct() %>%  # remove duplicated rows
#   select(-which(nearZero_info$nzv)) %>% # remove near zero variance variables
#   remove_empty() %>%  # removes empty rows and columns
#   remove_constant(na.rm = TRUE) %>%  # removes constsant columns
#   mutate(VarFactor = factor(VarFactor)) %>%  # categorical col as factor
#   mutate(VarDate = as.Date(VarDate, "%d-%m-%Y")) %>%  # convert chr to date 
```


## remove anomalies

- 

```{r}
# df <- df %>% 
#   filter(!((VarCategorical == 'Cat1')|
#            (VarCategorical == 'Cat2')|
#            (VarCategorical == 'Cat3'))) %>%  # remove specific categories
#   filter(!(VarNumerical > 120000)) %>%  # remove outlying values
#   select(-which(nearZero_info$nzv)) %>% # remove near zero variance variables
```


## additional variables

- 

```{r}
# new variable: sum of all VarWithPattern*
# df <- df %>%
#   rowwise %>%
#   mutate(AcceptedCmpAll = sum(c_across(contains("VarWithPattern"))))

```


```{r}
# check results from cleaning and added variables
head(df)
```
```{r}
summary(df)
```


## save processed data

```{r}
# df %>% write_csv(file = '../data/marketing_campaign_clean.csv')
```

